\begin{thebibliography}{99}

\bibitem{Cherry} SR Cherry, JA Sorenson, ME Phelps. ``Physics in
  Nuclear Medicine'', Saunders, 2012. ISBN 978-1-4160-5198-5.

%\bibitem{Sorenson87} JA Sorenson, ME Phelps. ``Physics in Nuclear Medicine'',
%     WB Saunders Company, Philadelphia, 1987.

\bibitem{Aanbevelingen} HY Oei, JAK Blokland, Nederlandse Vereniging voor
  Nucleaire Geneeskunde, ``Aanbevelingen Nucleaire Geneeskundige Diagnostiek'',
  Eburon, Delf, ISBN 90-5166-358-7.

\bibitem{Leerboek} JAJ Camps, MJPG van Kroonenburgh, P van Urk, KS
  Wiarda. ``Leerboek Nucleaire Geneeskunde'', Elsevier, ISBN 90 352
  2108 7.

\bibitem{Barrett} Harrison H.\ Barrett and Kyle J. Myers. "Foundations
  of image science." 1584 pages. ISBN 0-471-15300-1. Wiley-VCH,
  October 2003. 1 (2003).

\bibitem{IAEA} : IAEA book ``Nuclear Medicine Physics: A Handbook for
  Teachers and Students'', IAEA Reference Number: STI/PUB/1617,
  released on the internet on 2015-01-14, ISBN: 978-92-0-143810-2.\\
  \href{http://www-pub.iaea.org/books/IAEABooks/10368/Nuclear-Medicine-Physics-A-Handbook-for-Teachers-and-Students}
  {http://www-pub.iaea.org/books/IAEABooks/10368/Nuclear-Medicine-Physics-A-Handbook-for-Teachers-and-Students}

\bibitem{Stabin2008} M.G. Stabin, M.A. Emmons, M.J. Fernald,
  A.B. Brill, W.P. Segars. ``Use of realistic anthropomorphic models
  for calculation of radiation dose in nuclear Medicine''.
  International congress of the International Radiation Protection
  Association (IRPA), Argentina, 2008.
  
\bibitem{Haenscheid2013} H. H\"anscheid, Heribert, et al. ``EANM Dosimetry
  Committee series on standard operational procedures for
  pre-therapeutic dosimetry II. Dosimetry prior to radioiodine therapy
  of benign thyroid diseases.'' {\em European journal of nuclear medicine
  and molecular imaging} 40 (2013): 1126-1134.
  
\bibitem{Gaze2015} M.N. Gaze, et al. ``Feasibility of dosimetry-based
  high-dose 131I-meta-iodobenzylguanidine with topotecan as a
  radiosensitizer in children with metastatic neuroblastoma''.
  {\em Cancer biotherapy \& radiopharmaceuticals}, 2005, 20.2: 195-199.

\bibitem{Sangro2012} B. Sangro, M. I\~narrairaegui, J.I. Bilbao.
  ``Radioembolization for hepatocellular carcinoma'',
  {\em Journal of hepatology} 2012, 56.2: 464-473

\bibitem{Martin2006} Martin, James E. ``Physics for radiation protection: a
      handbook''. John Wiley \& Sons, 2006.
  
  %% \bibitem{Cantone} MC Cantone, C Hoeschen eds, ``Radiation Physics for
%%   Nuclear Medicine'', 2011. Springer. 
%%   \href{http://link.springer.com/book/10.1007/978-3-642-11327-7}
%%        {http://link.springer.com/book/10.1007/978-3-642-11327-7}

%% \bibitem{Stabin} Michael G. Stabin, ``Fundamentals of Nuclear Medicine
%%   Dosimetry'', 2008, Springer.
%%   \href{http://link.springer.com/book/10.1007/978-0-387-74579-4}
%%        {http://link.springer.com/book/10.1007/978-0-387-74579-4}

%% \bibitem{Cherry2006} S.R. Cherry, M. Dahlbom, M.E. Phelps, 
%%  ``PET, Physics, Instrumentation, and Scanners'', 2006, Springer.
%%   \href{http://link.springer.com/book/10.1007/0-387-34946-4}
%%        {http://link.springer.com/book/10.1007/0-387-34946-4}

%% \bibitem{Khalil}MM Khalil edt, ``Basic Sciences of Nuclear Medicine'',
%%   2011, Springer.
%%   \href{http://link.springer.com/book/10.1007/978-3-540-85962-8}
%%        {http://link.springer.com/book/10.1007/978-3-540-85962-8}

%% \bibitem{Zaidi} H.Zaidi edt, ``Quantitative Analysis in Nuclear
%%   Medicine Imaging'', 2006, Springer.
%%   \href{http://link.springer.com/book/10.1007/b107410}
%%        {http://link.springer.com/book/10.1007/b107410}

\end{thebibliography}
